Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization

### 2023-10-23 manuscript.Rmd
Update manuscript.Rmd
cite streamlined p hacking

Second, available resources limit theoretical risk.
Increasing theoretical risk $P(\neg{}E|\neg{}H)$ will usually decrease detectability $P(E|H)$ unless more resources are invested.
This is similar to the well known tradeoff between type-I error rate and statistical power.
Tasking preregistration with an increase in theoretical risk makes it difficult to balance this trade-off.
Mindlessly maximizing theoretical risk would either never produce evidence or require huge amounts of resources.
As noted before, we strive for high detectability and high theoretical risk.
Maximizing one at the expense of the other is not necessarily beneficial for increasing epistemic value but depends on the specific function they apply to judge evidence and their specific location on the curve.
One advantage of our framework is that researchers can employ it to balance the trade-off more effectively assuming they are willing to make some simplifiying assumptions.

@@

In case 1, there is no theoretical risk ($P(\neg{}E|\neg{}H) = 0$).
If we know that the results will be engineered to support the hypothesis no matter what, there is no reason to collect data.
Case 2 has a similar problem.
After all, the hypothesis that it had to happen the way it did happen is irrefutable.
In fact, both cases should be problematic to anyone who subscribes to the statistical relevancy condition because if we choose the hypothesis in accordance with the data or vice versa, without restrictions, they are not related anymore (i.e., observing the data does not tell us anything about the hypothesis and the other way around).
Case 3, is different since here the theoretical risk is not necessarily low but simply uncertain (and perhaps best represented by the dotted line in Figure \ref{fig:posterior}/\ref{fig:other-measures}).
In case 4, the theoretical risk is neither zero (unless the researcher plans to do run variations of analyses until a favourable outcome is obtained, then we have a particular instance case of 1) nor high (as this is the nature of exploratory approaches).
However, we can take advantage of computational reproducibility, use statistical properties, simulation or resampling methods, together with scientific reasoning, to get a reasonably certain evaluation of the theoretical risk and hence are in a somewhat favourable position (i.e., close to the solid line in n Figure \ref{fig:posterior}/\ref{fig:other-measures}).
This favorable position leads us to recommend preregistration of exploratory studies.
Case 5 shares the neither zero nor high theoretical risk of case 4 but has additional uncertainty about how much exploration was going on (how hard did researcher try to come up with favourable results).
Its low theoretical risk combined with high uncertainty about it, make it difficult to produce compelling evidence.

>>>>

Second, available resources limit theoretical risk.
Increasing theoretical risk $P(\neg{}E|\neg{}H)$ will usually decrease detectability $P(E|H)$ unless more resources are invested.
This is similar to the well known tradeoff between type-I error rate and statistical power.
Tasking preregistration with an increase in theoretical risk makes it difficult to balance this trade-off.
Mindlessly maximizing theoretical risk would either never produce evidence or require huge amounts of resources.
As noted before, we strive for high detectability and high theoretical risk in planning, conducting, and analyzing studies.
Maximizing one at the expense of the other is not necessarily beneficial for increasing epistemic value but depends on the specific function they apply to judge evidence and their specific location on the curve.
One advantage of our framework is that researchers can employ it to balance the trade-off more effectively assuming they are willing to make some simplifiying assumptions.

@@

In case 1, there is no theoretical risk ($P(\neg{}E|\neg{}H) = 0$).
If we know that the results will be engineered to support the hypothesis no matter what, there is no reason to collect data.
A prime example of this case is the $p_\text{pointless}$ metric [@husseyMethodStreamlinePhacking2021].
Case 2 has a similar problem.
After all, the hypothesis that it had to happen the way it did happen is irrefutable.
In fact, both cases should be problematic to anyone who subscribes to the statistical relevancy condition because if we choose the hypothesis in accordance with the data or vice versa, without restrictions, they are not related anymore (i.e., observing the data does not tell us anything about the hypothesis and the other way around).
Case 3, is different since here the theoretical risk is not necessarily low but simply uncertain (and perhaps best represented by the dotted line in Figure \ref{fig:posterior}/\ref{fig:other-measures}).
In case 4, the theoretical risk is neither zero (unless the researcher plans to do run variations of analyses until a favourable outcome is obtained, then we have a particular instance case of 1) nor high (as this is the nature of exploratory approaches).
However, we can take advantage of computational reproducibility, use statistical properties, simulation or resampling methods, together with scientific reasoning, to get a reasonably certain evaluation of the theoretical risk and hence are in a somewhat favourable position (i.e., close to the solid line in n Figure \ref{fig:posterior}/\ref{fig:other-measures}).
This favorable position leads us to recommend preregistration of exploratory studies.
Case 5 shares the neither zero nor high theoretical risk of case 4 but has additional uncertainty about how much exploration was going on (how hard did researcher try to come up with favourable results).
Its low theoretical risk combined with high uncertainty about it, make it difficult to produce compelling evidence.
